---
title: "Bernoulli Parameters in Example in JSM paper"
author: "Daniel Emaasit"
date: "March 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Bernoulli
This is an example in Section 5.5 of Gelman et al (2003), which studied coaching effects from eight schools. For simplicity, we call this example "eight schools."

In this model, we let `theta` be transformed parameters of `mu` and `eta` instead of directly declaring `theta` as parameters. By parameterizing this way, the sampler will run more efficiently. Assuming we have `8schools.stan` file in our working directory, we can prepare the data and fit the model as the following R code shows.

```{r data_and_fitModel}
bern_dat <- list(N = 10, y = c(0,1,0,0,0,0,0,0,0,1))

fit <- stan(file = 'bernoulli.stan', data = bern_dat, 
            iter = 1000, chains = 4, save_dso = TRUE)
```

## Model Results
We can also specify a Stan model using a character string by using argument `model_code` of function stan instead. However, this is not recommended.

The object `fit`, returned from function stan is an S4 object of class `stanfit`. Methods such as `print`, `plot`, and `pairs` are associated with the fitted result so we can use the following code to check out the results in fit.   

* `print` provides a summary for the parameter of the model as well as the log-posterior with name lp__ (see the following example output).   
* For more methods and details of class `stanfit`, see the help of class stanfit.

In particular, we can use `extract` function on `stanfit` objects to obtain the samples. extract extracts samples from the stanfit object as a list of arrays for parameters of interest, or just an array. In addition, S3 functions `as.array` and `as.matrix` are defined for `stanfit` object (using `help("as.array.stanfit")` to check out the help document in R).
```{r model_results}
print(fit)
plot(fit)
pairs(fit, pars = c("mu", "tau", "lp__"))

la <- extract(fit, permuted = TRUE) # return a list of arrays 
mu <- la$mu 

### return an array of three dimensions: iterations, chains, parameters 
a <- extract(fit, permuted = FALSE) 

### use S3 functions as.array (or as.matrix) on stanfit objects
a2 <- as.array(fit)
m <- as.matrix(fit)

## extract the posterior groups in a format that ggmcmc likes
library(ggmcmc)
samples <- ggmcmc::ggs(fit)

## show the histogram of sampled parameter values
ggmcmc::ggmcmc(D = samples, file = NULL, plot = 'ggs_histogram')

## look at the traceplot (should look like white noise)
ggmcmc::ggmcmc(D = samples, file = NULL, plot = 'ggs_traceplot')

## look at the full vs partial-density (should look the same)
ggmcmc::ggmcmc(D = samples, file = NULL, plot = 'ggs_compare_partial')

## look at the autocorrelations
ggmcmc::ggmcmc(D = samples, file = NULL, plot = 'ggs_autocorrelation')
```